EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-31 05_examples.tex
Updates from Overleaf

\section{Examples}\label{SecExamples}
In this section we illustrate and benchmark the functionalities and of the \NAME
library and its interfaces as a solver for DMFT through a variety of examples. We also discuss in detail the relevant code snippets from different programming languages. 
Some tasks -- particularly those associated with the implementation of the DMFT self-consistency -- are inherently independent of \NAME itself, and the impurity solver is agnostic about them. 
In the implementation using Fortran, Python or other \NAME API we adopt a reverse communication strategy, which requires the user to independently implement these parts or, equivalently, to exploit existing utilities. 
In our examples based on the Fortran APIs  we rely on external open-source libraries, such as DMFTtools\footnote{\url{https://github.com/aamaricci/DMFTtools}} (also developed by some of the authors) to perform specific tasks, including the evaluation of the local Green's function, the implementation of the self-consistency, the construction of the tight-binding model or the calculation of the kinetic energy. 
Where appropriate, we include comments that reference these external procedures.
On the other hand, the results obtained with W2Dynamics rely on the internal framework provided by the software itself -- see the script \texttt{DMFT.py} \cite{Wallerberger2019CPC}, which handles all these tasks seamlessly.
Similarly, the TRIQS library offers an extensive toolbox for the manipulation of Green’s functions and related quantities~\cite{Parcollet2015CPC}, making these tasks easily accessible while maintaining very high standards of optimization and accuracy.

\subsection{Hubbard model on the Bethe lattice (Fortran API, {\tt ed\_mode=normal})}\label{SecExamplesBetheDMFT}
The test bed for any DMFT implementation is conventionally considered 
the description of the Mott or metal-to-insulator transition (MIT) 
in a Bethe lattice, i.e., a Cayley tree with coordination number 
$z\to\infty$~\cite{Georges1996RMP}.
Here, we present a guided implementation of the DMFT 
solution for the Bethe lattice at $T=0$ using \NAME as impurity
solver.
The model under consideration is the Fermi-Hubbard Hamiltonian:
$$
H = -t \sum_{\langle ij\rangle,\s} c^\dagger_{i\s} c_{j\s} + 
    U \sum_i n_{i\up}n_{i\dw},
$$
where $c^\dagger_{i\s}$ ($c_{i\s}$) are the creation (annihilation) 
operators for an electron at site $i$ with spin $\s$, and 
$n_{i\s} = c^\dagger_{i\s} c_{i\s}$ is the corresponding occupation 
operator. The first sum runs over nearest-neighbor pairs
$\langle ij \rangle$.

The system is defined on a Bethe lattice with density of states
$\rho(\e)=\frac{1}{2D}\sqrt{D^2-\e^2}$,
where $D=2t$ is the half-bandwidth.
In DMFT, this lattice model is mapped onto a quantum impurity problem with
an effective electronic bath that must be determined
self-consistently.

Below, we present the key components of a basic DMFT implementation 
using \NAME for the Bethe lattice, employing either the Fortran or 
C++ APIs. Fully operational versions of the two codes can be found 
in the {\tt examples} directory of the \NAME source code. Both 
samples share a similar initial structure, including memory 
allocation, Bethe lattice creation, and solver initialization.

\setbox0=\hbox{%
  \begin{minipage}{0.465\linewidth}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK
   USE SCIFOR
   implicit none
   integer                :: Le=1000
   real(8)                :: wmixing=0.5d0
   real(8)                :: D=1d0
   integer                :: Nb
   real(8),allocatable    :: Bath(:)
   complex(8),allocatable :: Hloc(:,:,:,:)
   real(8),allocatable    :: Ebands(:),Dbands(:)
   complex(8),allocatable :: Smats(:,:,:,:,:)
   complex(8),allocatable :: Delta(:,:,:,:,:)
   !...  
   !Read variables
   call ed_read_input('inputED.conf')
   !Solver-specific arays
   allocate(Smats(Nspin,Nspin,Norb,Norb,Lmats))
   allocate(Delta(Nspin,Nspin,Norb,Norb,Lmats))
   allocate(Hloc(Nspin,Nspin,Norb,Norb))
   Hloc=0d0
   !...
   !Initialize Bethe arrays
   allocate(Ebands(Le),Dbands(Le))
   !From SciFortran:
   Ebands = linspace(-D,D,Le,mesh=de)
   Dbands = dens_bethe(Ebands,Wband)*de
   !Set local H
   call ed_set_hloc(Hloc)
\end{lstlisting}
\end{minipage}
}
\savestack{\listingA}{\box0}

\setbox0=\hbox{%
  \begin{minipage}{0.505\linewidth}
\begin{lstlisting}[style=cstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
#include <edipack_cbindings.h>
using namespace std;
//...
//Main variables
int Le = 1000;
int iloop = 0;
double wmixing = 0.5;
double D = 1.0;
//Read  variables    
char input[] = "inputED.conf"; 
read_input(input);      
//Dimensions
int64_t d[4] = {Nspin,Nspin,Norb,Norb};
int total_size = d[0] * d[1] * d[2] * d[3];
int total_size_n5 = total_size * Lmats;    
//Solver-specific arrays
vector<complex<double>> Hloc(total_size);
vector<complex<double>> Smats(total_size_n5);
vector<complex<double>> Delta(total_size_n5);
//...
//Initialize Bethe arrays
vector<double> Ebands, Dbands;
//Locally defined functions:
Ebands = linspace(-D,D,Le);
Dbands = dens_bethe(Ebands,Wband,de);
//set local H
ed_set_Hloc_single_N4(Hloc.data(), d);
\end{lstlisting}
\end{minipage}
}
\savestack{\listingB}{\box0}

\begin{tabular}{c|c}\label{list3}
{\bf Fortran} & {\bf C++}\\
  \stackinset{l}{}{t}{}{}{\listingA} & \stackinset{l}{}{t}{}{}{\listingB} 
\end{tabular}

The bath is described by the (unknown) function $\GG^{-1}_0$, called
the Weiss field ({\tt Weiss}). In the ED approach implemented in \NAME, 
the bath is approximated using a finite number of discrete energy levels. 
The function $\GG^{-1}_0$ is then used to determine the bath parameters 
$\vec{x} = \{V, h\}$ through the optimization method outlined in 
\secu{sSecFit}.

@@

\setbox0=\hbox{%
  \begin{minipage}{0.465\linewidth}
    \center
    \begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
   !Initialize solver
   Nb=ed_get_bath_dimension()
   allocate(bath(Nb))
   call ed_init_solver(bath)
\end{lstlisting}
\end{minipage}
}
\savestack{\listingA}{\box0}

\setbox0=\hbox{%
  \begin{minipage}{0.505\linewidth}
    \center
\begin{lstlisting}[style=cstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
// Initialize solver
int Nb;
vector<double> Bath(Nb);
Nb = get_bath_dimension_direct();
int64_t bath_dim[1] = {Nb};
init_solver_site(Bath.data(), bath_dim);
\end{lstlisting}
\end{minipage}
}
\savestack{\listingB}{\box0}

\begin{tabular}{c|c}\label{list4}
  \stackinset{l}{}{t}{}{}{\listingA} & \stackinset{l}{}{t}{}{}{\listingB} 
\end{tabular}

The iterative algorithm to solve the DMFT problem is based on
a reverse communication strategy and proceeds as follows:
\begin{itemize}

\item[{\tiny {\bf EDIpack}}] Call the exact diagonalization {\bf impurity solver}
  whose only input is the set of parameters $\vec{x}$ contained in a
  rank-1 array. All  \NAME options are controlled through input file
  specifications.

\item[{\tiny {\bf EDIpack}}] Retrieve the self-energy functions $\Sigma_{\a\b\s\s'}(i\omega)$ on the
  Matsubara axis using dedicated function provided by \NAME API.

\item[{\tiny {\it User}}] Evaluate the local interacting Green's function
$$
G_\mathrm{loc}(i\omega) = \int_{-D}^D \frac{\rho(\e)}{\zeta -\e}d\e
$$ 
with  $\zeta=i\omega+\mu-h^0-\Sigma(i\omega)$.

\item[{\tiny {\it User}}] Update the Weiss field via the {\bf self-consistency}
  relation: $\GG^{-1}_0(i\omega) = G_\mathrm{loc}(i\omega) +
    \Sigma(i\omega)$. For the Bethe lattice, this simplifies to
    $\Delta = \tfrac{D}{4}G_\mathrm{loc}$.

  \item[{\tiny {\it User} \textgreater\ {\bf EDIpack}}] Optimize the bath parameters $\vec{x}$ against the updated
    Weiss field,  potentially using the conjugate gradient procedures 
supplied by \NAME. Then, restart at step 1.
\end{itemize}

The first two steps are handled directly by \NAME routines, while during the subsequent steps the 
 control returns to the user, who must implement the algebraic updates required to 
close the self-consistency loop and optimize the bath.
Given the 
critical importance of this optimization, \NAME provides access to 
well-tested functions, ensuring stability and reproducibility of 
results. Alternative optimization methods can also be employed as 
needed. 
An example of implementation is provided in the following listings.

\setbox0=\hbox{%
  \begin{minipage}{0.46\linewidth}
    \center
    \begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
    !DMFT loop
do while(.not.converged.AND.iloop<nloop)
    iloop=iloop+1     
    !Call ED Solver
    call ed_solve(bath)     
    !Retrieve $\Sigma(i\omega_n)$
    call ed_get_sigma(Smats,'m')
    !Build local Green's function
    wfreq = pi/beta*(2*arange(1,Lmats)-1)
    do i=1,Lmats
       zeta= xi*wfreq(i)+xmu - Smats(1,1,1,1,i)
       Gmats(1,1,1,1,i)=sum(DOS/(zeta-Ene))*de  
    enddo
    !Self-consistency
    Delta = D/4d0*Gmats
    !Fitting -> new bath
    call ed_chi2_fitgf(Weiss,bath,ispin=1)     
    !Check convergence: from $\mathrm{DMFTtools}$
    converged=check_convergence(Delta,&
          dmft_error,Nsuccess,Nloop)
enddo
\end{lstlisting}
\end{minipage}
}
\savestack{\listingA}{\box0}

\setbox0=\hbox{%
  \begin{minipage}{0.51\linewidth}
    \center
\begin{lstlisting}[style=cstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
//DMFT loop
while (iloop < Nloop && !converged) {
  //Call ED Solver
  solve_site(Bath.data(),bath_dim,1,1);      
  //Retrieve $\Sigma(i\omega_n)$
  get_sigma_site_n5(Smats.data(),//
      0,0,wm.data(),Lmats,0);
  //Build a local Green's function
  for (int i=0;i<Lmats;++i) {
    zeta= wm[i] + xmu - Smats[i];
    Gmats[i] = complex<double>(0.0,0.0);
    for (int j=0; j< Le; j++) {
      Gmats[i]+=Dbands[j]/(zeta-Ebands[j]);
    }
  }
  //Self-consistency
  for (int i = 0; i < Lmats; ++i) {
    Delta[i] = 0.25 * D * Gmats[i];
  }
  //Fit -> new bath
  chi2_fitgf_single_normal_n5(Delta.data(),//
      delta_dim,Bath.data(),bath_dim,1,0,1);
  //Check convergence: local functions
  converged = check_convergence(Delta,//
      dmft_error, Nsuccess, Nloop, comm);  
}
\end{lstlisting}
\end{minipage}
}
\savestack{\listingB}{\box0}

\begin{tabular}{c|c}\label{list5}
  \stackinset{l}{}{t}{}{}{\listingA} & \stackinset{l}{}{t}{}{}{\listingB} 
\end{tabular}

\paragraph{Results.}
In the following, we present results for the interaction-driven Mott 
transition in the Bethe lattice, which captures the gradual 
transformation of a partially-filled metallic state into a correlated 
insulator.
To illustrate this, we report in panel (A) the evolution of the spectral 
function $-{\rm Im}G_\mathrm{loc}(\omega)/\pi$ as a function of the 
interaction strength $U$. Despite the inherently {\it spiky} nature 
of the spectrum, resulting from the finite number of poles in the 
discretized effective bath, the characteristic features of the Mott 
transition are clearly visible. At low energies, a renormalized 
quasi-particle peak develops at the Fermi level ($\omega = 0$). 
Simultaneously, the system exhibits the formation of incoherent 
high-energy features, which eventually evolve into well-defined 
Hubbard bands in the Mott insulating phase for $U > U_\mathrm{c}$, 
with $U_\mathrm{c} \simeq 2.8D$.

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{figures/figBethe.pdf}
    \caption{\label{figEx1}%
      \textbf{The Metal-insulator Mott transition.}
      (a) Evolution of the spectral function $-\Im{G}(\omega)/\pi$ as
      a function of increasing interaction $U$. The critical
      interaction $U_\mathrm{c}\simeq 2.8D$ separates the correlated metal $U<U_\mathrm{c}$ from
      the Mott insulator $U>U_\mathrm{c}$.
      (b)-(c) The corresponding evolution of the Matsubara self-energy
      $\Im\Sigma(i\omega)$ (b) and
      real-axis one $\Re\Sigma(\omega)$ across the Mott
      transition. For a particle-hole symmetric case, both allow to
      estimate the renormalization constant $Z$ (see main text), using
      linear order expansion in frequency (blue solid lines). The
      values of $Z$ are reported in the legend.
      The Mott insulating solution is associated with a singularity at
      $\omega=0$ of $\Im\Sigma$.
    }
\end{figure}

The formation of a spectral gap separating the Hubbard bands in the 
Mott insulating phase is associated with the divergence of the 
imaginary part of the self-energy at the Fermi level. This divergence 
reflects the complete localization of the electrons, effectively 
suppressing coherent quasi-particle excitations. Specifically, 
causality dictates that the real part of the self-energy must also 
grow significantly near the singularity, making it impossible to 
satisfy the quasi-particle pole equation
$$
\omega+\mu-h^0-\e-\Re{\Sigma}(\omega)=0,
$$
which governs the formation of coherent excitations near the Fermi 
level.

Panels (B) and (C) illustrate this phenomenon by showing the evolution 
of the self-energy $\Sigma$. In panel (B), we present the Matsubara 
self-energy ${\rm Im}\Sigma(i\omega)$ in the low-energy regime. As the 
interaction strength $U$ increases, this function progressively 
grows, eventually diverging as the critical interaction threshold 
$U > U_\mathrm{c}$ is crossed.
This divergence along the Matsubara axis is directly linked to the
particle-hole symmetry of the Bethe lattice, which pins the  ${\rm
  Im}\Sigma$ singularity at $\omega = 0$.

Panel (C) complements this picture by displaying the real part 
$\Re{\Sigma}(\omega)$ in the same low-energy regime. Here, increasing 
$U$ leads to a rapid rise of this component, culminating in a 
discontinuous behavior as the critical point is approached. This 
discontinuity directly reflects the divergence in the imaginary part
on the real-axis,  confirming the transition to the Mott insulating state.

@@

The linear fits highlighted in panels (B) and (C), along with the 
corresponding $Z$ values provided in the legends, clearly indicate that 
the slope of the self-energy at low energy increases with $U$ on both 
the Matsubara and real axes. At the transition point, this slope 
diverges, reflecting the onset of complete electron localization as 
$Z \to 0$, consistent with the divergence 
${\rm Im}\Sigma(\omega) \rightarrow -\infty$.

\subsubsection{Finite temperature solution (w2dynamics interface, CT-HYB benchmark)}\label{SecExamplesBetheDMFTw2d}
To demonstrate the usage of the w2dynamics interface to \NAME,  here we briefly discuss how to solve the same problem, i.e. the Hubbard model on the Bethe lattice within DMFT using w2dynamics. 
In order to showcase the capabilities of \NAME to address low-temperature problems, we compare the continuous-time Quantum Monte Carlo (CTQMC) solver using the hybridization expansion method (CT-HYB) included in w2dynamics against the \NAME solver at finite temperature.

In contrast to the \NAME API that only enables access to the impurity solver and bath optimization procedures, but requires the user to implement (using different methods) the DMFT algorithm themselves, the standard way to use w2dynamics is fundamentally different: a single Python script {\tt DMFT.py} is provided that takes care of the whole DMFT calculation via dedicated classes to implement generic self-consistency. The w2dynamics calculation is then entirely controlled by a model-dependent parameters file {\tt Parameters.in}, which contains a number of variable specifications including options to control the ED solver inherited from \NAME. Further information about the functioning of w2dynamics can be found in Ref.~\onlinecite{Wallerberger2019CPC}

Another important difference concerns the initial point of the iterative DMFT solution algorithm: while \NAME starts from a given discrete bath, w2dynamics is initialized with a zero self-energy function or alternatively reads this quantity from a converged solution file. When using the \NAME ED interface, the initial Weiss field is determined using self-consistency and finally an initial discrete bath is obtained through the \NAME bath optimization procedure.

>>>>

\section{Examples}\label{SecExamples}
In this section we illustrate and benchmark the functionalities and of the \NAME
library and its interfaces as a solver for DMFT through a variety of examples. We also discuss in detail the relevant code parts from different programming languages.

Some tasks -- particularly those associated with the implementation of the DMFT self-consistency -- are inherently independent of \NAME itself, and the impurity solver is agnostic about them. 
Thus, in the implementations using Fortran, Python or other \NAME interfaces to scientific toolboxes we adopt a reverse communication strategy, which requires the user to independently carry out these parts or, equivalently, to exploit existing utilities.

In our examples based on the Fortran APIs we rely on external open-source libraries, such as DMFTtools\footnote{\url{https://github.com/aamaricci/DMFTtools}} (also developed by some of the authors) to perform specific tasks, including the evaluation of the local Green's function, the implementation of the self-consistency, the construction of the tight-binding model or the calculation of the kinetic energy. 
Where appropriate, we include comments that reference these external procedures.
On the other hand, the results obtained with W2Dynamics rely on the internal framework provided by the software itself -- see the script \texttt{DMFT.py} \cite{Wallerberger2019CPC}, which handles all these tasks seamlessly.
Similarly, the TRIQS library offers an extensive framework for the manipulation of Green’s functions and related quantities~\cite{Parcollet2015CPC}, making these tasks easily accessible while maintaining very high standards of optimization and accuracy.

\subsection{Hubbard model on the Bethe lattice (Fortran API, {\tt ed\_mode=normal})}\label{SecExamplesBetheDMFT}
The test bed for any DMFT implementation is conventionally considered 
the description of the Mott or metal-to-insulator transition (MIT) 
in a Bethe lattice, i.e., a Cayley tree with coordination number 
$z\to\infty$~\cite{Zhang1993PRLb,Caffarel1994PRL,Georges1996RMP,Rozenberg1999PRL,Kotliar1999EPJB,Kotliar2000PRL,Kotliar2002PRL}.
Here, we present a guided implementation of the DMFT 
solution for the Bethe lattice at $T=0$ using \NAME as impurity
solver.
The model under consideration is the Fermi-Hubbard Hamiltonian:
$$
H = -t \sum_{\langle ij\rangle,\s} c^\dagger_{i\s} c_{j\s} + 
    U \sum_i n_{i\up}n_{i\dw},
$$
where $c^\dagger_{i\s}$ ($c_{i\s}$) are the creation (annihilation) 
operators for an electron at site $i$ with spin $\s$, and 
$n_{i\s} = c^\dagger_{i\s} c_{i\s}$ is the corresponding occupation 
operator. The first sum runs over nearest-neighbor pairs
$\langle ij \rangle$.
We consider the system defined on a Bethe lattice with density of states
$\rho(\e)=\frac{2}{\pi D^2}\sqrt{D^2-\e^2}$,
where $D=2t$ is the half-bandwidth.
Within DMFT framework~\cite{Georges1996RMP}, this lattice model is mapped onto a quantum impurity problem with an effective electronic bath that must be determined self-consistently.

Below, we present the key components of a basic DMFT implementation 
using \NAME for the Bethe lattice, employing either the Fortran or 
C++ APIs. Fully operational versions of the two codes can be found 
in the {\tt examples} directory of the \NAME source code. Both 
samples share a similar initial structure, including memory 
allocation, creation of the Bethe DOS and solver initialization.

\begin{center}
\begin{minipage}[t]{0.49\linewidth}
\textbf{Fortran}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK
   USE SCIFOR
   implicit none
   integer                :: Le=1000
   real(8)                :: wmixing=0.5d0
   real(8)                :: D=1d0
   integer                :: Nb
   real(8),allocatable    :: Bath(:)
   complex(8),allocatable :: Hloc(:,:,:,:)
   real(8),allocatable    :: Ebands(:),Dbands(:)
   complex(8),allocatable :: Smats(:,:,:,:,:)
   complex(8),allocatable :: Delta(:,:,:,:,:)
   !...  
   !Read variables
   call ed_read_input('inputED.conf')
   !Solver-specific arays
   allocate(Smats(Nspin,Nspin,Norb,Norb,Lmats))
   allocate(Delta(Nspin,Nspin,Norb,Norb,Lmats))
   allocate(Hloc(Nspin,Nspin,Norb,Norb))
   Hloc=0d0
   !...
   !Initialize Bethe arrays
   allocate(Ebands(Le),Dbands(Le))
   !From SciFortran:
   Ebands = linspace(-D,D,Le,mesh=de)
   Dbands = dens_bethe(Ebands,Wband)*de
   !Set local H
   call ed_set_hloc(Hloc)
\end{lstlisting}
\end{minipage}
%
\begin{minipage}[t]{0.49\linewidth}
\textbf{C++}
\begin{lstlisting}[style=cstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
#include <edipack_cbindings.h>
using namespace std;
//...
//Main variables
int Le = 1000;
int iloop = 0;
double wmixing = 0.5;
double D = 1.0;
//Read  variables    
char input[] = "inputED.conf"; 
read_input(input);      
//Dimensions
int64_t d[4] = {Nspin,Nspin,Norb,Norb};
int total_size = d[0] * d[1] * d[2] * d[3];
int total_size_n5 = total_size * Lmats;    
//Solver-specific arrays
vector<complex<double>> Hloc(total_size);
vector<complex<double>> Smats(total_size_n5);
vector<complex<double>> Delta(total_size_n5);
//...
//Initialize Bethe arrays
vector<double> Ebands, Dbands;
//Locally defined functions:
Ebands = linspace(-D,D,Le);
Dbands = dens_bethe(Ebands,Wband,de);
//set local H
ed_set_Hloc_single_N4(Hloc.data(), d);
$\phantom{.}$
$\phantom{.}$
\end{lstlisting}
\end{minipage}
\end{center}

The bath is described by the (unknown) function $\GG^{-1}_0$, i.e. 
the Weiss field ({\tt Weiss}). In the ED method implemented in \NAME, 
the bath is approximated using a finite number of discrete energy levels. 
The Weiss field $\GG^{-1}_0$ is then used to determine the bath parameters $\vec{x} = \{\hat{V}, \hat{h}\}$ through the optimization method outlined in \secu{sSecFit}.

@@

\begin{center} 
\begin{minipage}[t]{0.49\linewidth}
\textbf{Fortran}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
   !Initialize solver
   Nb=ed_get_bath_dimension()
   allocate(bath(Nb))
   call ed_init_solver(bath)
$\phantom{.}$   
$\phantom{.}$   
\end{lstlisting}
\end{minipage}
%
\begin{minipage}[t]{0.49\linewidth}
\textbf{C++}
\begin{lstlisting}[style=cstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
// Initialize solver
int Nb;
vector<double> Bath(Nb);
Nb = get_bath_dimension_direct();
int64_t bath_dim[1] = {Nb};
init_solver_site(Bath.data(), bath_dim);
\end{lstlisting}
\end{minipage}
\end{center}

The iterative algorithm to solve the DMFT problem proceeds as follows:
\begin{itemize}  
\item[{\tiny {\bf EDIpack}}] Call the \NAME {\bf impurity solver}
  whose only input is the set of parameters $\vec{x}$ contained in a
  rank-1 array. All  \NAME options are controlled through input file
  specifications.

\item[{\tiny {\bf EDIpack}}] Retrieve the self-energy functions $\Sigma_{\a\b\s\s'}(i\omega)$ on the
  Matsubara axis using dedicated function {\tt ed\_get\_sigma} available in the \NAME API.

\item[{\tiny {\it User}}] Evaluate the local interacting Green's function
$$
G_\mathrm{loc}(i\omega) = \int_{-D}^D \frac{\rho(\e)}{\zeta -\e}d\e
$$ 
with  $\zeta=i\omega+\mu-h^0-\Sigma(i\omega)$. Note that this step can be performed analytically for the Bethe lattice~\cite{Georges1996RMP} or completely substituted by retrieving the impurity Green's function $G_\mathrm{imp}$ via the {\tt ed\_get\_gimp} function.

\item[{\tiny {\it User}}] Update the Weiss field via the {\bf self-consistency}
  relation: $\GG^{-1}_0(i\omega) = G_\mathrm{loc}(i\omega) +
    \Sigma(i\omega)$. For the Bethe lattice, this simplifies to
    $\Delta = \tfrac{D}{4}G_\mathrm{loc}$ or, equivalently, $\Delta = \tfrac{D}{4}G_\mathrm{imp}$.

  \item[{\tiny {\it User} \textgreater\ {\bf EDIpack}}] Optimize the bath parameters $\vec{x}$ against the updated
    Weiss field using the conjugate gradient procedures supplied by \NAME. Then, restart at step 1.
\end{itemize}

The first two steps are handled directly by \NAME routines, while during the subsequent steps the 
 control returns to the user, who must implement the algebraic updates required to 
close the self-consistency loop and optimize the bath.
Given the 
critical importance of this optimization, \NAME provides access to 
well-tested functions, ensuring stability and reproducibility of 
results. Alternative optimization methods can also be employed as 
needed~\cite{Mejuto-Zaera2020PRB,Huang2023PRB,Nakatsukasa2018SJOSC}. 
An example of implementation is provided in the following listings.

\begin{center} 
\begin{minipage}[t]{0.49\linewidth}
\textbf{Fortran}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
!DMFT loop
do while(.not.converged.AND.iloop<nloop)
    iloop=iloop+1     
    !Call ED Solver
    call ed_solve(bath)     
    !Retrieve ${\color{comment-color} \Sigma(i\omega_n)}$
    call ed_get_sigma(Smats,'m')
    !Build local Green's function
    wfreq = pi/beta*(2*arange(1,Lmats)-1)
    do i=1,Lmats
       zeta= xi*wfreq(i)+xmu - Smats(1,1,1,1,i)
       Gmats(1,1,1,1,i)=sum(DOS/(zeta-Ene))*de  
    enddo

    !Self-consistency
    Delta = 0.25d0*D*Gmats

    !Fitting -> new bath
    call ed_chi2_fitgf(Weiss,bath,ispin=1)

    !Check convergence: from ${\color{comment-color}\mathrm{DMFTtools}}$
    converged=check_convergence(Delta,&
          dmft_error,Nsuccess,Nloop)

enddo
$\phantom{.}$   
\end{lstlisting}
\end{minipage}
%
\begin{minipage}[t]{0.49\linewidth}
\textbf{C++}
\begin{lstlisting}[style=cstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
//DMFT loop
while (iloop < Nloop && !converged) {
  //Call ED Solver
  solve_site(Bath.data(),bath_dim,1,1);      
  //Retrieve ${\color{comment-color} \Sigma(i\omega_n)}$
  get_sigma_site_n5(Smats.data(),//
      0,0,wm.data(),Lmats,0);
  //Build a local Green's function
  for (int i=0;i<Lmats;++i) {
    zeta= wm[i] + xmu - Smats[i];
    Gmats[i] = complex<double>(0.0,0.0);
    for (int j=0; j< Le; j++) {
      Gmats[i]+=Dbands[j]/(zeta-Ebands[j]);
    }
  }
  //Self-consistency
  for (int i = 0; i < Lmats; ++i) {
    Delta[i] = 0.25 * D * Gmats[i];
  }
  //Fit -> new bath
  chi2_fitgf_single_normal_n5(Delta.data(),//
      delta_dim,Bath.data(),bath_dim,1,0,1);
  //Check convergence: local functions
  converged = check_convergence(Delta,//
      dmft_error, Nsuccess, Nloop, comm);  
}
\end{lstlisting}
\end{minipage}
\end{center}

\paragraph{Results.}
In the following, we present \NAME results for the interaction-driven MIT obtained with previous implementations. The MIT captures the gradual transformation of a partially-filled metallic state into a correlated insulator.

To illustrate this, we report in panel (A) the evolution of the spectral function $-{\rm Im}G_\mathrm{loc}(\omega)/\pi$ as a function of the  interaction strength $U$. 
Despite the inherently {\it spiky} nature of the spectrum, resulting from the finite number of poles in the 
discretized effective bath, the characteristic features of the Mott 
transition are clearly visible. At low energies, a renormalized 
quasi-particle peak develops at the Fermi level ($\omega = 0$). 
Simultaneously, the system exhibits the formation of incoherent 
high-energy features, which eventually evolve into well-defined 
Hubbard bands in the Mott insulating phase for $U > U_\mathrm{c}$, 
with $U_\mathrm{c} \simeq 2.8D$.

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{figures/figBethe.pdf}
    \caption{\label{figEx1}%
      \textbf{The metal-insulator Mott transition.}
      (a) Evolution of the spectral function $-\Im{G}(\omega)/\pi$ as
      a function of increasing interaction $U$. The critical
      interaction $U_\mathrm{c}\simeq 2.8D$ separates the correlated metal $U<U_\mathrm{c}$ from
      the Mott insulator $U>U_\mathrm{c}$.
      (b)-(c) The corresponding evolution of the Matsubara self-energy
      $\Im\Sigma(i\omega)$ (b) and
      real-axis one $\Re\Sigma(\omega)$ across the Mott
      transition. For a particle-hole symmetric case, both allow to
      estimate the renormalization constant $Z$ (see main text), using
      linear order expansion in frequency (blue solid lines). The
      values of $Z$ are reported in the legend.
      The Mott insulating solution is associated with a singularity at
      $\omega=0$ of $\Im\Sigma$.
    }
\end{figure}

The formation of a spectral gap separating the Hubbard bands in the 
Mott insulating phase is associated with the divergence of the 
imaginary part of the self-energy at the Fermi level. This divergence 
reflects the complete localization of the electrons, effectively 
suppressing coherent quasi-particle excitations. Causality dictates that the real part of the self-energy must also 
grow significantly near the singularity, making it impossible to 
satisfy the quasi-particle pole equation
$$
\omega+\mu-h^0-\e-\Re{\Sigma}(\omega)=0,
$$
which governs the formation of coherent excitations near the Fermi 
level.
Panels (B) and (C) illustrate this phenomenon by showing the evolution of the self-energy $\Sigma$. In panel (B), we present the Matsubara self-energy ${\rm Im}\Sigma(i\omega)$ in the low-energy regime. As the  interaction strength $U$ increases, this function progressively grows, eventually diverging as the critical interaction threshold $U > U_\mathrm{c}$ is crossed.
This divergence along the Matsubara axis is directly linked to the
particle-hole symmetry of the Bethe lattice, which pins the  ${\rm
  Im}\Sigma$ singularity at $\omega = 0$.
Panel (C) complements this picture by displaying the real part 
$\Re{\Sigma}(\omega)$ on the real-axis near the Fermi level. Here, increasing $U$ leads to a rapid rise of this component, culminating in a discontinuous behavior as the critical point is approached. This 
discontinuity directly reflects the divergence in the imaginary part
on the real-axis, confirming the transition to the Mott insulating state.

@@

The linear fits highlighted in panels (B) and (C), along with the 
corresponding $Z$ values provided in the legends, clearly indicate that 
the slope of the self-energy at low energy increases with $U$ on both 
the Matsubara and real axes. At the transition point, this slope 
diverges, reflecting the onset of complete electron localization as 
$Z \to 0$, consistent with the singular behavior  
$-{\rm Im}\Sigma(\omega\to0) \to\infty$.

\subsubsection{Finite temperature solution (W2Dynamics interface, CT-HYB benchmark)}\label{SecExamplesBetheDMFTW2D}
To demonstrate the usage of the W2Dynamics interface to \NAME,  here we briefly discuss how to solve the same problem, i.e. the Hubbard model on the Bethe lattice within DMFT using W2Dynamics. 
In order to showcase the capabilities of \NAME to address low-temperature problems, we compare the continuous-time Quantum Monte Carlo (CTQMC) solver using the hybridization expansion method (CT-HYB) included in W2Dynamics against the \NAME solver at finite temperature.

Unlike \NAME, which provides only the impurity solver and bath optimization procedures and requires the user to implement the DMFT algorithm themselves (potentially using various methods), W2Dynamics adopts a fundamentally different approach: a single Python script, {\tt DMFT.py}, handles the entire DMFT calculation, leveraging dedicated classes that implement the generic self-consistency. The W2Dynamics calculation is then entirely controlled by a model-dependent parameters file {\tt Parameters.in}, which contains a number of variable specifications including options to control the ED solver inherited from \NAME. Further information about the functioning of W2Dynamics can be found in Ref.~\cite{Wallerberger2019CPC}

Another important difference concerns the initial point of the iterative DMFT solution algorithm: while \NAME starts from a given discrete bath, W2Dynamics is initialized with a zero self-energy function or alternatively reads this quantity from a converged solution file. Thus, when using the \NAME interface in W2Dynamics, the initial Weiss field is determined using self-consistency and  an initial discrete bath is obtained through the \NAME bath optimization procedure.

The W2Dynamics configuration file for the solution of the Bethe lattice problem at finite temperature reads: 
\begin{lstlisting}[style=mybash,language={},numbers=none,basicstyle={\scriptsize\ttfamily}]
[General]
DOS             = Bethe           # support for the Bethe lattice is built-in
half-bandwidth  = 1               # list of half-bandwidths per orbital
NAt             = 1               # number of impurities
beta            = 100             # inverse temperature
mu              = 1.0             # chemical potential set to achieve half-filling
EPSN            = 0.0             # turns off filling-based chemical potential search
DMFTsteps       = 100             # given no convergence checking, we might want fewer
magnetism       = para            # symmetrize self-energies per spin
FileNamePrefix  = bethe_dmft_U2   # prefix for the output file name
fileold         = bethe_dmft*hdf5 # file to read an initial self-energy from
readold         = 0               # iteration number to read initial self-energy from, 0 turns off
mixing          = 0.5             # mixing, but mixes self-energies and not Weiss fields
mixing_strategy = linear          # linear mixing as in the Fortran example
FTType          = none            # (for CT-HYB): use the NFFT-measured G
solver          = EDIPACK         # use EDIpack as impurity solver, not default CTHYB
